---
title: "PEcAn API tutorial"
author: "Alexey Shiklomanov, Rob Kooper"
date: "2021-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Pretty-print long IDs
options(scipen = 12, pillar.sigfig = 12)
```

NOTE: Disconnect from VPN before trying this!
Rob's servers look very suspicious to corporate VPNs.

These are Rob's PEcAn test servers.

```{r}
ip1 <- "ZZZ.ZZZ.ZZZ.ZZZ"
```

The full URL is `http://<IP address>`.

```{r}
library(rpecanapi)
url <- sprintf("http://%s", ip1)
print(url)
```

Connect to the server and test the connection.
Normally, there would be authentication here, but these test servers have no authentication.

```{r}
server <- connect(url)
ping(server)
get.status(server)
```

Here, we see what models are available.
This endpoint doesn't have an R frontend yet, so we access it directly using `httr`.

```{r}
library(httr)
library(dplyr)

models <- GET(file.path(server$url, "api", "availableModels/")) %>%
  content() %>%
  bind_rows()

models
```

For our first test, we'll use SIPNET.
We can extract its model ID from the `models` query output above. 

```{r}
sipnet_id <- models %>%
  filter(model_name == "SIPNET") %>%
  pull(model_id)

get.model(server, sipnet_id)
```

We can use the API `search.sites` function to search for sites.

```{r}
search.sites(server, "Harvard")
```

I happen to know the Harvard Forest ID, so we'll just use that here.
`get.site` tells us a bit more about the site.

```{r}
# Harvard Forest
site_id <- 646
get.site(server, site_id)
```

Now, we have everything we need to submit the workflow.
Note that this runs fairly quickly because the MERRA data are already downloaded.

```{r}
wf <- submit.workflow(
  server, sipnet_id, site_id,
  pfts = "temperate.coniferous",
  start_date = "2012-06-01",
  end_date = "2012-08-01",
  inputs = list(met = list(source = "MERRA"))
)
wf
```

Use the URL below to track progress:

```{r}
print(file.path(
  server$url,
  "pecan",
  sprintf("05-running.php?workflowid=%s&hostname=docker", wf$workflow_id)
))
```

Once the run has finished, we can check its status and look at some outputs.

```{r}
rpecanapi::get.workflow.status(server, wf$workflow_id)

runs <- rpecanapi::get.runs(server, wf$workflow_id)
runid <- runs$runs$id
rpecanapi::plot_run_vars(server, runid, 2012, "GPP")
```

We can also download the outputs directly for more sophisticated analysis.

```{r}
rpecanapi::download.run.output(server, runid, "2012.nc")

library(ncdf4)
nc <- nc_open("2012.nc")
gpp <- ncvar_get(nc, "GPP")
npp <- ncvar_get(nc, "NPP")
resp <- ncvar_get(nc, "TotalResp")
dtime <- ncvar_get(nc, "time")
time <- as.POSIXct("2020-01-01 00:00:00", tz="UTC") + as.difftime(dtime, units = "days")

clrs <- c("green4", "brown", "blue")
matplot(time, cbind(gpp, resp, npp), type = "l", lty = "solid", col = clrs,
        xlab = "Time", ylab = expression("Flux" ~ (kg * C ~ m^-2 ~ s^-1)))
abline(h = 0, lty = "dashed", col = "black")
legend("topleft", c("GPP", "TotResp", "NPP"), lty = "solid", col = clrs)
```

Let's try a more complex example with ED2.
This requires more configuration because ED2 requires a lot more inputs.
Note that the argument `inputs` targets the inputs section of the XML,
while `workflow_list_mods` can be used to arbitrarily modify _any_ part of the XML
(based on base R's `modifyList` function).
(In principle, you could provide nonsense to all the other arguments and 
pass all information using the `workflow_list_mods` argument).

```{r}
ed_id <- models %>%
  filter(model_name == "ED2.2") %>%
  pull(model_id)

wf2 <- submit.workflow(
  server, ed_id, site_id,
  pfts = "temperate.Early_Hardwood",
  start_date = "2012-06-01",
  end_date = "2012-06-30",
  # Passed directly into the `inputs` tag
  inputs = list(met = list(source = "MERRA"),
                lu = list(id = 294),
                thsum = list(id = 295),
                veg = list(id = 296),
                soil = list(id = 297)),
  # Passed to `modifyList` to modify the list representation of the workflow XML
  workflow_list_mods = list(model = list(
    phenol.scheme = 0,
    ed_misc = list(output_month = 12),
    edin = "ED2IN.r2.2.0"
  ))
)
wf2
```

Again, follow along using this URL.

```{r}
print(file.path(
  server$url,
  "pecan",
  sprintf("05-running.php?workflowid=%s&hostname=docker", wf2$workflow_id)
))
```

ED2 post processing doesn't fully work.

```{r}
runs <- rpecanapi::get.runs(server, wf2$workflow_id)
runid <- runs$runs$id
rpecanapi::download.run.output(server, runid, "analysis-T-2012-00-00-000000-g01.h5")
```

```{r}
nc2 <- nc_open("analysis-T-2012-00-00-000000-g01.h5")
gpp <- ncvar_get(nc2, "FMEAN_GPP_PY")
time <- as.POSIXct("2012-06-01 00:00:00", "UTC") + as.difftime(seq_along(gpp), units = "hours")
ii <- time < "2012-07-01"
plot(time[ii], gpp[ii], type = "l")
```
